\section{Methods}
\label{sec:methods}

All models share the same high-level structure: an \emph{encoder} $f_\theta$ that maps
each raw sequence to a fixed-length representation, a \emph{fusion} module that combines
the enhancer and promoter representations into a joint feature vector, and a
\emph{classification head} that maps the joint vector to a scalar interaction probability.
\Cref{fig:overview} sketches this pipeline.

\begin{figure}[H]
\centering
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[
    node distance = 0.9cm and 1.5cm,
    box/.style    = {rectangle, draw, rounded corners, minimum width=2.2cm,
                     minimum height=0.8cm, align=center, font=\small},
    arrow/.style  = {-Latex, thick},
  ]
  \node[box, fill=blue!10]  (en_seq)  {Enhancer\\Sequence};
  \node[box, fill=blue!10,  below=of en_seq] (pr_seq) {Promoter\\Sequence};
  \node[box, fill=orange!20, right=of en_seq] (en_enc) {Encoder\\$f_\theta$};
  \node[box, fill=orange!20, right=of pr_seq] (pr_enc) {Encoder\\$f_\theta$};
  \node[box, fill=green!20,  right=2.0cm of en_enc, yshift=-0.45cm]
        (fusion) {Fusion\\Module};
  \node[box, fill=red!15,    right=of fusion] (head) {Classif.\\Head};
  \node[box, fill=gray!15,   right=of head]   (out)  {$\hat{y}$};
  \draw[arrow] (en_seq) -- (en_enc) node[midway, above, font=\scriptsize]{$s_e$};
  \draw[arrow] (pr_seq) -- (pr_enc) node[midway, above, font=\scriptsize]{$s_p$};
  \draw[arrow] (en_enc.east) -- ++(0.4,0) |- (fusion.west)
               node[near end, above, font=\scriptsize]{$\mathbf{h}_e$};
  \draw[arrow] (pr_enc.east) -- ++(0.4,0) |- (fusion.west)
               node[near end, below, font=\scriptsize]{$\mathbf{h}_p$};
  \draw[arrow] (fusion) -- (head) node[midway, above, font=\scriptsize]{$\mathbf{m}$};
  \draw[arrow] (head)   -- (out);
\end{tikzpicture}}%
\caption{High-level architecture of DeepChrInteract-v2.
         Encoder weights may be shared (parameter tying) or independent (dual-branch).}
\label{fig:overview}
\end{figure}

We organise the fourteen encoder variants into five groups:
(A) classical CNN reproductions,
(B) RNN/LSTM family,
(C) Transformer family,
(D) linear recurrence and SSM,
(E) self-supervised and pre-trained encoders.

% ============================================================
\subsection{Group A: Classical CNN (Reproduction)}
% ============================================================

\subsubsection{M1: CNN Single-Branch}

One-hot input $\mathbf{X} \in \mathbb{R}^{5 \times L}$ is processed by two stages
of strided Conv1d (kernel 24, stride 4), each followed by Batch Normalisation~\cite{batch_norm},
ReLU, and max-pooling.
Channel sizes are 64 in the first stage and 128 in the second.
The output is globally average-pooled and passed through two Linear--Dropout(0.5) layers
($d = 512$).
This reproduces \texttt{model\_onehot\_cnn\_one\_branch} from the original codebase.

\subsubsection{M2: CNN Dual-Branch}

Two independent M1 encoders process the enhancer and promoter sequences separately.
Their output vectors $\mathbf{h}_e, \mathbf{h}_p \in \mathbb{R}^{512}$ are passed to
the fusion module.
This reproduces \texttt{model\_onehot\_cnn\_two\_branch}.

\subsubsection{M3: \texorpdfstring{$k$}{k}-mer Embedding + CNN}

The 6-mer tokenised sequence $\mathbf{t} \in \mathbb{Z}^{L-5}$ is embedded via
$\mathbf{E} \in \mathbb{R}^{4097 \times 128}$ (optionally initialised with DNA2Vec
weights) and processed by four Conv1d--ReLU--MaxPool blocks.
Reproduces \texttt{model\_embedding\_cnn\_two\_branch}.

% ============================================================
\subsection{Group B: RNN / LSTM Family}
% ============================================================

\subsubsection{M4: Bidirectional LSTM}

Two-layer BiLSTM~\cite{lstm} ($d_h = 256$ per direction, dropout 0.3) processes the
one-hot sequence in both directions; the concatenated final hidden states form
$\mathbf{h}_e \in \mathbb{R}^{512}$.

\subsubsection{M5: mLSTM (xLSTM)}

The matrix-memory LSTM of Beck~et~al.~\cite{xlstm} replaces the scalar LSTM cell
$c_t$ with a matrix $\mathbf{C}_t \in \mathbb{R}^{d_k \times d_v}$, updated via
\begin{equation}
  \mathbf{C}_t = f_t \cdot \mathbf{C}_{t-1} + i_t \cdot \mathbf{v}_t \mathbf{k}_t^\top,\quad
  \mathbf{h}_t = \mathbf{o}_t \odot \tfrac{\mathbf{C}_t \mathbf{q}_t}
    {\max(\|\mathbf{C}_t \mathbf{q}_t\|_\infty, 1)},
\end{equation}
with stabilised normalisation preventing gradient explosion.
We stack two bidirectional mLSTM layers ($d_k = d_v = 128$).

% ============================================================
\subsection{Group C: Transformer Family}
% ============================================================

\subsubsection{M6: Transformer Encoder (Standard)}

A strided CNN front-end (stride 16) downsamples $L$ to $\approx 625$ tokens before
sinusoidal positional encoding and four Transformer encoder
layers~\cite{attention_all_you_need}
($d_{\mathrm{model}} = 256$, $n_{\mathrm{head}} = 8$, $d_{\mathrm{ff}} = 1024$,
dropout 0.1).
A prepended \texttt{[CLS]} token is used as the sequence representation.

\subsubsection{M7: Linear Transformer}

Katharopoulos~et~al.~\cite{linear_transformer} approximate softmax attention via a
kernel feature map $\phi$:
\begin{equation}
  \mathrm{Attn}_i = \frac{\phi(\mathbf{q}_i)^\top
    \textstyle\sum_{j}\phi(\mathbf{k}_j)\mathbf{v}_j^\top}
  {\phi(\mathbf{q}_i)^\top\sum_{j}\phi(\mathbf{k}_j)},
\end{equation}
reducing complexity from $O(L^2 d)$ to $O(Ld^2)$.
No CNN front-end is needed: the full 10 kbp sequence is processed at
single-nucleotide resolution ($\phi = \mathrm{ELU}(\cdot)+1$).

\subsubsection{M8: iTransformer}

Attention is applied across the five \emph{channels}
($\mathbf{X} \in \mathbb{R}^{C \times L}$, $C = 5$) rather than across positions,
yielding $O(C^2 L)$ complexity ($C = 5 \ll L$)~\cite{itransformer}.
The feed-forward network captures position-wise nonlinear patterns per channel.
This captures nucleotide co-occurrence statistics invisible to positional attention.

% ============================================================
\subsection{Group D: Linear Recurrence and SSM}
% ============================================================

\subsubsection{M9: Mamba}

Mamba~\cite{mamba} evolves an input-dependent hidden state:
\begin{equation}
  \mathbf{h}_t = \overline{\mathbf{A}}(x_t)\,\mathbf{h}_{t-1}
               + \overline{\mathbf{B}}(x_t)\,x_t,\quad
  y_t = \mathbf{C}(x_t)\,\mathbf{h}_t,
\end{equation}
where $\overline{\mathbf{A}}, \overline{\mathbf{B}}, \mathbf{C}$ depend on $x_t$,
enabling selective focus on functional motifs.
Four Mamba blocks ($d_{\mathrm{model}} = 256$) with mean pooling; no CNN front-end.

\subsubsection{M10: RWKV}

Time-mixing propagates context via exponential position decay:
\begin{equation}
  \mathbf{o}_t = \mathbf{W}_o \cdot \sigma(r_t) \odot
    \frac{\sum_{s \leq t} e^{-(t-s)w+k_s} v_s}
         {\sum_{s \leq t} e^{-(t-s)w+k_s}},
\end{equation}
where $w$ is a learnable per-channel decay rate~\cite{rwkv}.
Training is parallelised via cumulative sum; inference is $O(1)$ per step.
Four RWKV blocks ($d_{\mathrm{model}} = 256$) with mean pooling.

% ============================================================
\subsection{Group E: Hybrid and Pre-trained Encoders}
% ============================================================

\subsubsection{M11: CNN + BiLSTM}

The CNN front-end (M1, truncated after pooling, output $\approx 500$ tokens) serves
as a motif detector; the BiLSTM (M4) captures long-range dependencies over the
resulting feature map.

\subsubsection{M12: CNN + Transformer}

Identical to M11 but replaces the BiLSTM with the Transformer encoder (M6).
The CNN downsamples the sequence to manageable length before self-attention.

\subsubsection{M13: DNA LLM Encoder}

Pre-trained DNA foundation models are used as frozen or fine-tuned feature
extractors (see \Cref{sec:data}, Encoding Strategy C):
DNABERT~\cite{dnabert}, DNABERT-2~\cite{dnabert2},
Nucleotide Transformer~\cite{nucleotide_transformer},
and HyenaDNA~\cite{hyenadna}.
A linear projection aligns the LLM output dimension to the common head size.

\subsubsection{M14: MAE Pre-trained Transformer}

The Transformer encoder (M6 architecture) is first pre-trained on \emph{unlabelled}
DNA sequences via a Masked Autoencoder~\cite{mae} objective: a high fraction
($p = 0.75$) of nucleotide tokens are masked, and the decoder reconstructs the
original one-hot vectors.
After pre-training, the decoder is discarded and the encoder is fine-tuned on the
labelled EPI pairs.
Compared to LLM-based encoders, this approach allows self-supervised adaptation
specifically to the target cell-type data distribution without requiring
large-scale external corpora.

% ============================================================
\subsection{Fusion Module}
\label{sec:fusion}
% ============================================================

Given $\mathbf{h}_e, \mathbf{h}_p \in \mathbb{R}^d$, the fusion module produces
$\mathbf{m}$ according to one of the strategies in \Cref{tab:fusion}.

\begin{table}[H]
\centering
\caption{Supported fusion strategies. $d_{\mathrm{out}}$: dimension fed to the head.}
\label{tab:fusion}
\begin{tabular}{llcc}
\toprule
Strategy & Formula & $d_{\mathrm{out}}$ & Extra params \\
\midrule
Concat        & $[\mathbf{h}_e;\mathbf{h}_p]$
              & $2d$ & 0 \\
Add           & $\mathbf{h}_e + \mathbf{h}_p$
              & $d$  & 0 \\
Subtract      & $\mathbf{h}_e - \mathbf{h}_p$
              & $d$  & 0 \\
Multiply      & $\mathbf{h}_e \odot \mathbf{h}_p$
              & $d$  & 0 \\
Bilinear      & $\mathbf{h}_e^\top \mathbf{W}_b \mathbf{h}_p$
              & $1$  & $d^2$ \\
Concat+Sub+Mul & $[\mathbf{h}_e;\mathbf{h}_p;\mathbf{h}_e{-}\mathbf{h}_p;\mathbf{h}_e{\odot}\mathbf{h}_p]$
              & $4d$ & 0 \\
\bottomrule
\end{tabular}
\end{table}

\textsc{Concat+Sub+Mul} is the recommended default: it is parameter-free and captures
both directional differences (subtraction) and co-activation patterns (multiplication).

\subsection{Classification Head}

All models share a two-layer MLP:
\begin{equation}
  \hat{y} = \sigma\!\bigl(\mathbf{W}_2\,\mathrm{ReLU}(\mathbf{W}_1 \mathbf{m}
            + \mathbf{b}_1) + \mathbf{b}_2\bigr),
\end{equation}
trained with binary cross-entropy and Adam~\cite{adam}
($\beta_1 = 0.9$, $\beta_2 = 0.999$, weight decay $10^{-5}$).
